---
title: "DSDA 1995"
subtitle: "Data Science and Society using R"
author: "Lecture 17 | Shiny in Quarto"
format: html 
embed-resources: true
server: shiny
layout: sidebar
---

```{r}
#| include: FALSE
library(tidyverse)
knitr::opts_chunk$set(eval = TRUE, echo = TRUE)
```

## Plan for Today

-   Explore Shiny apps in R
-   Explore basic UI
-   Explore basic reactivity

<br>
<br>
<br>
<br>
<br>
<br>

----

## Shiny App Introduction

-   What is Shiny?

    -   Shiny is a framework for creating web applications using R code

    -   It is designed primarily with data scientists in mind, and to that end, you can create pretty complicated Shiny apps with no knowledge of HTML, CSS, or JavaScript

    -   On the other hand, Shiny doesn’t limit you to creating trivial or prefabricated apps: its user interface components can be easily customized or extended, and its server uses reactive programming to let you create any type of back end logic you want

-   What do people use Shiny for?

    -   Create dashboards that track important high-level performance indicators, while facilitating drill down into metrics that need more investigation

    -   Replace hundreds of pages of PDFs with interactive apps that allow the user to jump to the exact slice of the results that they care about

    -   Communicate complex models to a non-technical audience with informative visualizations and interactive sensitivity analysis

<br>
<br>
<br>
<br>
<br>
<br>

----

### Shiny App

-   Let's install `shiny` and start putting together a dashboard

-   We are also going to install `DT` (replaces Shiny functions for data tables)

```{r}
#install.packages("rmarkdown")
#install.packages("shiny")
#install.packages("DT")


library(shiny)
library(DT)
```

<br>
<br>
<br>
<br>
<br>
<br>

----

### Shiny

- With RMarkdown

  -   There are several ways to create a Shiny app. The simplest is to create a new directory for your app, and put a single file called `app.R` in it

  -   This `app.R` file will be used to tell Shiny both how your app should look, and how it should behave
  
- With Quarto

  - Quarto doesn't require `fluidPage()` or `shinyApp()` — your document is the UI. Code chunks containing `render*()` functions are reactive and act like Shiny server code.

```{r}

library(shiny)

"Hello, world!"
```

<br>
<br>
<br>
<br>
<br>
<br>

----

### Adding UI Controls and Behavior

-   Let's add some inputs and outputs to our app

```{r}
selectInput("dataset", "Dataset", choices = ls("package:datasets"))
```

-   New functions

    + `selectInput()` is an input control that lets the user interact with the app by providing a value. In this case, it’s a select box with the label “Dataset” and lets you choose one of the built-in datasets that come with R
    
-   Now let's define some server functions

    -   Shiny uses reactive programming to make apps interactive

    -   It involves telling Shiny how to perform a computation, not ordering Shiny to actually go do it

-   We’ll tell Shiny how to fill in the summary and table outputs in the sample app by providing the “recipes” for those outputs

- In the first output we will get the summary of the data. In the second, we will get the first six rows

<br>
<br>

```{r}
#| eval: false
#| context: server

output$summary <- renderPrint({
  dataset <- get(input$dataset, "package:datasets")
  summary(dataset)
})

output$headtable <- renderTable({
  dataset <- get(input$dataset, "package:datasets")
  head(dataset)
})
```

```{r}
#| context: server

output$summary <- renderPrint({
  dataset <- get(input$dataset, "package:datasets")
  summary(dataset)
})

output$headtable <- renderTable({
  dataset <- get(input$dataset, "package:datasets")
  head(dataset)
})
```

```{r}
#| echo: false
#| panel: center

### Summary of the dataset
verbatimTextOutput("summary")

### Head of the dataset
tableOutput("headtable")
```

<br>
<br>
<br>
<br>
<br>
<br>

----

## Basic UI

-   Let's examine some of the HTML inputs and outputs that Shiny provides

<br>
<br>
<br>
<br>
<br>
<br>

----

### Inputs

-   All input functions have the same first argument: `inputId`

    -   This is the identifier used to connect the front end with the back end

-   The `inputId` has two constraints

    -   It must be a simple string that contains only letters, numbers, and underscores (no spaces, dashes, periods, or other special characters allowed!). Name it like you would name a variable in R

    -   It must be unique. If it’s not unique, you’ll have no way to refer to this control in your server function

<br>
<br>

```{r}
#textInput(inputId, label, value = "", width = NULL, placeholder = NULL)

textInput("name", "Name", "What's your name?")
textInput("age", "Age", "What's your age?")
passwordInput("password", "What's your password?")
textAreaInput("story", "Tell me about yourself", rows = 3)
```

<br>
<br>

-   If you want to collect numeric values use the following

    -   `numericInput()`

    -   `sliderInput()`
<br>
<br>

```{r}
numericInput("num", "Number one", value = 0, min = 0, max = 100)
sliderInput("num2", "Number two", value = 50, min = 0, max = 100)
sliderInput("rng", "Range", value = c(10, 20), min = 0, max = 100)
```

<br>
<br>

-   You can have inputs that collect the following

    -   Dates (`dateInput()` or `dateRangeInput()`)

    -   limited choices (`selectInput()` or `radioButtons()`)

    -   Action buttons (`actionButton()`)

<br>
<br>
<br>
<br>
<br>
<br>

----

### Outputs

-   Outputs in the UI create placeholders that are later filled by the server function

-   Like inputs, outputs take a unique ID as their first argument: if your UI specification creates an output with ID "plot", you’ll access it in the server function with `output$plot`

-   Each output function on the front end is coupled with a render function in the back end

-   There are three main output types

    -   Text

    -   Tables

    -   Plots

-   Let's look at `text`

<br>
<br>

```{r}
#| eval: false
#| context: server

output$greeting <- renderText({
  paste("Hello,", input$name, "— Age:", input$age)
})
```


```{r}
#| context: server

output$greeting <- renderText({
  paste("Hello,", input$name, "— Age:", input$age)
})
```

```{r}
#| panel: fill

### creating a greeting
verbatimTextOutput("greeting")

```

<br>
<br>
<br>
<br>
<br>
<br>

----

-   Let's look at `tables`

```{r}
#| eval: false
#| context: server

output$headtable1 <- renderTable({
  head(mtcars)
})

#install.packages("DT")
#library(DT)
output$dttable <- DT::renderDT(mtcars,
             options = list(pageLength = 5))

```


<br>
<br>

```{r}
#| context: server

output$headtable1 <- renderTable({
  head(mtcars)
})

#install.packages("DT")
#library(DT)
output$dttable <- DT::renderDT(mtcars,
             options = list(pageLength = 5))

```

```{r}
#| echo: false
#| panel: center

### Head of the dataset
tableOutput("headtable1")

### Head of the dataset
DT::dataTableOutput("dttable")
```

<br>
<br>

-   `renderTable()` render a static table of data, showing all the data at once

    -   Is most useful for small, fixed summaries

-   `DT::renderDT()` and render a dynamic table, showing a fixed number of rows along with controls to change which rows are visible

    -   `DT::renderDT()` is most appropriate if you want to expose a complete data frame to the user

<br>
<br>
<br>
<br>
<br>
<br>

----

- Let's look at `plot`

- You can display any type of R graphic with `plotOutput()` and `renderPlot()`

```{r}
#| eval: false
#| context: server

output$plot <- renderPlot({
  plot(1:5)
}, res = 96)

output$plot1 <- renderPlot({
  ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy))
})

```

<br>
<br>

```{r}
#| context: server

library(tidyverse)

output$plot <- renderPlot({
  plot(1:5)
}, res = 96)


output$plot1 <- renderPlot({
  ggplot(data = mpg) + 
    geom_point(mapping = aes(x = displ, y = hwy))
})

```

```{r}
#| echo: true
#| panel: center

plotOutput("plot")

plotOutput("plot1")
```

<br>
<br>
<br>
<br>
<br>
<br>

----

## Basic Reactivity

- In Shiny, you express your server logic using reactive programming. 

- Reactive programming is an elegant and powerful programming paradigm, but it can be disorienting at first because it’s a very different paradigm to writing a script

  + The key idea of reactive programming is to specify a graph of dependencies so that when an input changes, all related outputs are automatically updated
  
<br>
<br>
<br>
<br>
<br>
<br>

----

### The Server Function

- The `ui` object that contains the HTML is presented to every user of your app 

- The `ui` is simple because every user gets the same HTML

- The `server` is more complicated because every user needs to get an independent version of the app; when user A moves a slider, user B shouldn’t see their outputs change

- To achieve this independence, Shiny invokes your `server()` function each time a new session starts

- Server functions take three parameters: `input`, `output`, and `session`

<br>
<br>
<br>
<br>
<br>
<br>

----

### Input

- The input argument is a list-like object that contains all the input data sent from the browser, named according to the input ID

```{r}
#| eval: false

numericInput("count", label = "Number of values", value = 100)

```

- Then you can access the value of that input with `input$count`

  + It will initially contain the value 100, and it will be automatically updated as the user changes the value in the browser
  
- Unlike a typical list, input objects are read-only. If you attempt to modify an input inside the server function, you’ll get an error  
  
```{r}
#| eval: false

input$count <- 10  

```

- This error occurs because input reflects what’s happening in the browser, and the browser is Shiny’s “single source of truth”

- If you could modify the value in R, you could introduce inconsistencies, where the input slider said one thing in the browser, and input$count said something different in R. That would make programming challenging



<br>
<br>
<br>
<br>
<br>
<br>

----

### Output

- `output` is very similar to `input`: it’s also a list-like object named according to the output ID

- The main difference is that you use it for sending output instead of receiving input 

```{r}
#| eval: false

output$greeting <- renderText({
  paste("Hello,", input$name, "— Age:", input$age)
})
```

- Render function

  + It sets up a special reactive context that automatically tracks what inputs the output uses
  
  + It converts the output of your R code into HTML suitable for display on a web page
  
- Like the `input`, the `output` is picky about how you use it. You’ll get an error if

  + You forget the render function
  
  + You attempt to read from an `output`



<br>
<br>
<br>
<br>
<br>
<br>

----

### Reactive Programming

- An app is going to be pretty boring if it only has inputs or only has outputs. The real magic of Shiny happens when you have an app with both

```{r}
textInput("name", "What's your name?")
```

```{r}
textOutput("greeting1")
```

```{r}
#| context: server

output$greeting1 <- renderText({
    paste0("Hello ", input$name, "!")
})

```

<br>
<br>

  
- This is the big idea in Shiny: you don’t need to tell an output when to update, because Shiny automatically figures it out for you

- How does it work?

  + Read this as “paste together ‘hello’ and the user’s name
  
  + Then send it to `output$greeting1`
  
    + But Shiny performs the action every time we update `input$name`
    
      + The app works because the code doesn’t tell Shiny to create the string and send it to the browser, but instead, it informs Shiny how it could create the string if it needs to


<br>
<br>
<br>
<br>
<br>
<br>

----

### Reactive Expressions

- Reactive expressions are important because they give Shiny more information so that it can do less recomputation when inputs change, making apps more efficient, and they make it easier for humans to understand the app by simplifying the reactive graph

- Reactive expressions have both `inputs` and `outputs`

```{r}
#| eval: false

freqpoly <- function(x1, x2, binwidth = 0.1, xlim = c(-3, 3)) {
  df <- data.frame(
    x = c(x1, x2),
    g = c(rep("x1", length(x1)), rep("x2", length(x2)))
  )

  ggplot(df, aes(x, colour = g)) +
    geom_freqpoly(binwidth = binwidth, size = 1) +
    coord_cartesian(xlim = xlim)
}

t_test <- function(x1, x2) {
  test <- t.test(x1, x2)
  
  # use sprintf() to format t.test() results compactly
  sprintf(
    "p value: %0.3f\n[%0.2f, %0.2f]",
    test$p.value, test$conf.int[1], test$conf.int[2]
  )
}

#simulated data
x1 <- rnorm(100, mean = 0, sd = 0.5)
x2 <- rnorm(200, mean = 0.15, sd = 0.9)

freqpoly(x1, x2)
cat(t_test(x1, x2))
```

<br>
<br>

- Let's wrap these simulations in in a Shiny app, where we can interactively tweak the inputs

<br>
<br>

```{r}
#| context: setup

freqpoly <- function(x1, x2, binwidth = 0.1, xlim = c(-3, 3)) {
  df <- data.frame(
    x = c(x1, x2),
    g = c(rep("x1", length(x1)), rep("x2", length(x2)))
  )
  ggplot(df, aes(x, colour = g)) +
    geom_freqpoly(binwidth = binwidth, size = 1) +
    coord_cartesian(xlim = xlim)
}

t_test <- function(x1, x2) {
  test <- t.test(x1, x2)
  sprintf("p value: %0.3f\n[%0.2f, %0.2f]", test$p.value, test$conf.int[1], test$conf.int[2])
}
```

::: {layout-ncol=3}

```{r}
numericInput("n1", label = "n (Dist 1)", value = 1000, min = 1)
numericInput("mean1", label = "µ (Dist 1)", value = 0, step = 0.1)
numericInput("sd1", label = "σ (Dist 1)", value = 0.5, min = 0.1, step = 0.1)
```

```{r}
numericInput("n2", label = "n (Dist 2)", value = 1000, min = 1)
numericInput("mean2", label = "µ (Dist 2)", value = 0, step = 0.1)
numericInput("sd2", label = "σ (Dist 2)", value = 0.5, min = 0.1, step = 0.1)
```

```{r}
numericInput("binwidth", label = "Bin width", value = 0.1, step = 0.1)
sliderInput("range", label = "Range", value = c(-3, 3), min = -5, max = 5)
```

:::

```{r}
#| context: server

output$plot2 <- renderPlot({
  x1 <- rnorm(input$n1, input$mean1, input$sd1)
  x2 <- rnorm(input$n2, input$mean2, input$sd2)
  freqpoly(x1, x2, binwidth = input$binwidth, xlim = input$range)
}, res = 96)


output$text1 <- renderText({
  x1 <- rnorm(input$n1, input$mean1, input$sd1)
  x2 <- rnorm(input$n2, input$mean2, input$sd2)
  t_test(x1, x2)
})
```

```{r}
textOutput("text1")
plotOutput("plot2")
```



<br>
<br>
<br>
<br>
<br>
<br>

----

## Tidy Evaluation

- Let's examine using `tidyverse` code within the Shiny environment

- Imagine we want to create an app that allows you to filter a numeric variable to select rows that are greater than a threshold

```{r}
# Load libraries and define variables
library(tidyverse)

num_vars <- c("carat", "depth", "table", "price", "x", "y", "z")

```


```{r}
selectInput("var", "Variable", choices = num_vars)
numericInput("min", "Minimum", value = 1)
```

```{r}
tableOutput("output")
```

```{r}
#| eval: false
#| context: server

data <- reactive({
  req(input$var, input$min)
  diamonds %>% filter(.data[[input$var]] > input$min)
})

output$output <- renderTable({
  head(data())
})
```

```{r}
#| context: server

data <- reactive({
  req(input$var, input$min)
  diamonds %>% filter(.data[[input$var]] > input$min)
})

output$output <- renderTable({
  head(data())
})
```

<br>
<br>

- What do we notice? 

  + Did the code run? 
  
  + Was the code executed correctly? 
  
- all the rows have values of `carat` less than 1

- `dplyr` thinks you have asked for `filter(diamonds, "carat" > 1)`

  + This is a problem of indirection
  
    + Normally when using tidyverse functions you type the name of the variable directly in the function call. But now you want to refer to it indirectly: the variable `(carat)` is stored inside another variable `(input$var)`
    
- Environment Variable

  + is a “programming” variables that you create with <-. input$var is an environment variable
  
- Data Variable

  +  is “statistical” variable that lives inside a data frame. `carat` is a data variable
  
- Our Problem:

  + We have a data variable (carat) stored inside an environment variable (input$var), and we need some way to tell dplyr this




<br>
<br>
<br>
<br>
<br>
<br>

----

## For Next Class

-   Review what you learned today

-   Do new set of readings:

    -   We are going to continue our discussion of Shiny
    -   [MSR Chapter 5]{style="color:#000E2F"}
    -   [MSR Chapter 6]{style="color:#000E2F"}
    -   Follow along the exercises with your own computer

-   Bring your laptops to class


