---
title: "DSDA 1995"
subtitle: "Data Science and Society using R"
author: "Lecture 15 | Geospatial Data I"
format: 
  revealjs:
    slide-number: c/t
    scrollable: true
    theme: [simple, custom.scss]
    embed-resources: true
editor: visual
---

```{r}
#| include: FALSE
library(tidyverse)
knitr::opts_chunk$set(eval = TRUE, echo = TRUE)
```

## Plan for Today

-   Explore plotting geolocated data in R
-   Examine plotting data with the `sp` and `sf` packages

## Types of Maps

-   We are going to create a few different types of maps, which will require different R packages

    -   Static Maps

        -   Most common type of visual output

            -   `plot()` most commonly used command from base are

            -   Can utilize `ggplot()` and `tmap()`

    -   Animated Maps

        -   Are becoming more widely used than in the past

            -   We will make these with `tmap()`

    -   Interactive Maps

        -   While static and animated maps can enliven geographic datasets, interactive maps can take them to a new level

            -   `leaflet()`

## Plotting Simple Features (`sf`)

-   As we have already briefly seen, the `sf` package extends the base plot command, so it can be used on `sf` objects

-   We are going to create a choropleth map with `sf`

```{r}
#| eval: TRUE
#| echo: TRUE
#| out-width: 100%
#| out-height: 100%

library(sf)

philly_crimes_sf <- st_read("PhillyCrimerate.shp")

glimpse(philly_crimes_sf)

plot(philly_crimes_sf)
```

-   Notice that the plot function will try and provide a map of every variable in the data set

## Plotting Simple Features (`sf`)

-   To plot a single attribute, we need to provide an object of class `sf`, like so

```{r}
#| eval: TRUE
#| echo: TRUE
#| out-width: 100%
#| out-height: 100%


plot(philly_crimes_sf$homic_rate) # this is a numeric vector!

plot(philly_crimes_sf["homic_rate"])
```

-   Based on the plot above, it would seem that we have a lot of "empty" geographic units or unevenly distributed

```{r}
#| eval: TRUE
#| echo: TRUE
#| out-width: 100%
#| out-height: 100%

ggplot(philly_crimes_sf) +
  geom_histogram(mapping = aes(x = homic_rate))
```

## Plotting Simple Features (`sf`)

-   Let's break the plot up into quantiles to better distinguish the census tracts with low values

```{r}
#| eval: TRUE
#| echo: TRUE
#| out-width: 100%
#| out-height: 100%

plot(philly_crimes_sf["homic_rate"], 
     main = "Philadelphia homicide density per square km", 
     breaks = "quantile")
```

## Mapping with `ggplot()`

-   `ggplot` is a widely used and powerful plotting library for R as we have utilized over the course of the semester

    -   It is not specifically geared towards mapping, but one can generate great maps

```{r}
#| eval: TRUE
#| echo: TRUE
#| out-width: 100%
#| out-height: 100%

ggplot(philly_crimes_sf) + 
  geom_sf(aes(fill=homic_rate))
```

## Mapping with `ggplot()`

-   Homicide rate is a continuous variable and is plotted by ggplot as such. If we wanted to plot our map as a ‘true’ choropleth map we need to convert our continuous variable into a categorical one, according to whichever brackets we want to use

```{r}
#| eval: TRUE
#| echo: TRUE
#| out-width: 100%
#| out-height: 100%

breaks_qt <- classInt::classIntervals(c(min(philly_crimes_sf$homic_rate) - .00001, philly_crimes_sf$homic_rate), n = 7, style = "quantile")

philly_crimes_sf <- mutate(philly_crimes_sf, homic_rate_cat = cut(homic_rate, breaks_qt$brks)) 

ggplot(philly_crimes_sf) + 
    geom_sf(aes(fill=homic_rate_cat)) +
    scale_fill_brewer(palette = "OrRd") 
```

## Mapping with `tmap()`

-   `tmap` is specifically designed to make creation of thematic maps more convenient

    -   Uses `ggplot` syntax

-   Useful things to know:

    -   `tm_shape()` where we provide the `sf` object (we could also provide an `SpatialPolygonsDataframe`)

    -   `tm_polygons()` where we set the attribute variable to map, the break style, and a title

## Mapping with `tmap()`

```{r}
#| eval: TRUE
#| echo: TRUE
#| out-width: 100%
#| out-height: 100%

library(tmap)

tm_shape(philly_crimes_sf) +
  tm_polygons("homic_rate", 
              style="quantile", 
              title="Philadelphia \nhomicide density \nper sqKm")
```

-   `tmap` has a very nice feature that allows us to give basic interactivity to the map. We can switch from “plot” mode into “view” mode and call the last plot, like so:

```{r}
#| eval: TRUE
#| echo: TRUE
#| out-width: 100%
#| out-height: 100%

tm_shape(philly_crimes_sf) +
  tm_polygons("homic_rate", 
              style="quantile", 
              title="Philadelphia \nhomicide density \nper sqKm")

tmap_mode("view")
tmap_last()
```

## Mapping with `leaflet()`

-   `leaflet` provides bindings to the ‘Leaflet’ JavaScript library, “the leading open-source JavaScript library for mobile-friendly interactive maps”

-   We have already seen a simple use of leaflet in the `tmap` example

-   Let’s build up the map step by step

```{r}
#| eval: TRUE
#| echo: TRUE
#| out-width: 100%
#| out-height: 100%


library(leaflet) 

#reproject
philly_WGS84 <- st_transform(philly_crimes_sf, 4326)

leaflet(philly_WGS84) %>%
  addPolygons()
```

## Mapping with `leaflet()`

-   To map the homicide density we use `addPolygons()`

```{r}
#| eval: TRUE
#| echo: TRUE
#| out-width: 100%
#| out-height: 100%


pal_fun <- colorQuantile("YlOrRd", NULL, n = 5)

p_popup <- paste0("<strong>Homicide Density: </strong>", philly_WGS84$homic_rate)

leaflet(philly_WGS84) %>%
  addPolygons(
    stroke = FALSE, # remove polygon borders
    fillColor = ~pal_fun(homic_rate), # set fill color with function from above and value
    fillOpacity = 0.8, smoothFactor = 0.5, # make it nicer
    popup = p_popup)  # add popup
```

## Mapping with `leaflet()`

-   We can add a basemap, which defaults to OSM, with `addTiles()`

```{r}
#| eval: TRUE
#| echo: TRUE
#| out-width: 100%
#| out-height: 100%


leaflet(philly_WGS84) %>%
  addPolygons(
    stroke = FALSE, 
    fillColor = ~pal_fun(homic_rate),
    fillOpacity = 0.8, smoothFactor = 0.5,
    popup = p_popup) %>%
  addTiles()
```

-   Lastly, we add a legend with `addLegend()`

```{r}
#| eval: TRUE
#| echo: TRUE
#| out-width: 100%
#| out-height: 100%


leaflet(philly_WGS84) %>%
  addPolygons(
    stroke = FALSE, 
    fillColor = ~pal_fun(homic_rate),
    fillOpacity = 0.8, smoothFactor = 0.5,
    popup = p_popup) %>%
  addTiles() %>%
  addLegend("bottomright",  # location
            pal=pal_fun,    # palette function
            values=~homic_rate,  # value to be passed to palette function
            title = 'Philadelphia homicide density per sqkm') # legend title
```

## For Next Class

-   Review what you learned today

-   Do new set of readings:

    -   We are going to continue or discuss of relational data
    -   [USDR Chapter 3.5-3.6]{style="color:#000E2F"}
    -   [GCR Chapter 8.3-8.6]{style="color:#000E2F"}
    -   [MSR Chapter 1-4]{style="color:#000E2F"}
    -   [MSR Chapter 12]{style="color:#000E2F"}
        -   Follow along the exercises with your own computer

-   Bring your laptops to class
