---
title: "DSDA 1995"
subtitle: "Data Science and Society using R"
author: "Lecture 8 | Data Transformation II"
format: 
  revealjs:
    slide-number: c/t
    scrollable: true
    theme: [simple, custom.scss]
    embed-resources: true
editor: visual
---

```{r}
#| include: FALSE
library(tidyverse)
knitr::opts_chunk$set(eval = TRUE, echo = TRUE)
```

## Plan for Today

-   Explore data transformation in R

::: extrapad
-   Explore the `dplyr` package/functions within `tidyverse`
:::

::: extrapad
-   In class activity
:::

## `mutate()`

-   `mutate()` allows you to manipulate the data and create new columns
    -   The new columns of data are added to the end of your data set

::: extrapad
-   **Helpful functions in `mutate()`**
    -   `mutate()` allows you to manipulate the data and create new columns

    -   `transmute()` mutates the data, but only keeps the new variables that you created
:::

------------------------------------------------------------------------

```{r}
#| eval: TRUE
#| echo: TRUE

#we are going to work with the flights data set
nycflights13::flights

flights <- nycflights13::flights
```

<br>

```{r}
#| eval: TRUE
#| echo: TRUE

#we are going to create a subset of the data to work with
flights_sml <- flights |>
  select(year:day, 
         ends_with("delay"),
         distance, 
         air_time)
```

<br>

```{r}
#| eval: TRUE
#| echo: TRUE

#creating two new variables from those already in the data set
flights_sml |>
  mutate(gain = dep_delay - arr_delay,
         speed = distance / air_time * 60)
```

<br>

```{r}
#| eval: TRUE
#| echo: TRUE

#you can refer to columns that you have just created
flights_sml |>
  mutate(gain = dep_delay - arr_delay,
         hours = air_time / 60,
         gain_per_hour = gain / hours)
```

------------------------------------------------------------------------

-   **`transmute()`**

```{r}
#| eval: TRUE
#| echo: TRUE

#use transmute if we only want to keep the manipulated variables
flights_sml |>
  transmute(gain = dep_delay - arr_delay,
            hours = air_time / 60,
            gain_per_hour = gain / hours)
```

## Useful Creation Functions

-   **Arithmetic Operators**
    -   `+`
    -   `-`
    -   `*`
    -   `/`
    -   `^`

------------------------------------------------------------------------

-   **Modular Arithmetic**
    -   `%/%` (integer division)
    -   `%%` (remainder)

```{r}
#| eval: TRUE
#| echo: TRUE

flights |> 
  transmute(dep_time,
            hour = dep_time %/% 100,
            minute = dep_time %% 100)
```

## `summarise()`

-   Collapses a data frame to a single row

```{r}
#| eval: TRUE
#| echo: TRUE

flights |>
  summarise(delay = mean(dep_delay, na.rm = TRUE))

mean(flights$dep_delay, na.rm = TRUE)
```

------------------------------------------------------------------------

-   Not terribly useful unless we pair it with `group_by()`
    -   Changes the unit of analysis from the complete dataset to individual groups
    -   When you use the dplyr verbs on a grouped data frame they’ll be automatically applied “by group”

<br>

```{r}
#| eval: TRUE
#| echo: TRUE

by_day <- flights |>
  group_by(year, month, day)

by_day |> 
  summarise(delay = mean(dep_delay, na.rm = TRUE))
```

<br>

-   Together `group_by()` and `summarise()` provide one of the tools that you’ll use most commonly when working with dplyr: grouped summaries

## Pipes `|>`

-   Pipes allow us to combine multiple operations together

```{r}
#| eval: TRUE
#| echo: TRUE

by_dest <- group_by(flights, dest)

delay <- summarise(by_dest,
  count = n(),
  dist = mean(distance, na.rm = TRUE),
  delay = mean(arr_delay, na.rm = TRUE)
)

delay <- filter(delay, count > 20, dest != "HNL")
```

<br>

**The Steps to Prepare the Data**

-   Group flights by destination
-   Summarize to compute distance, average delay, and number of flights
-   Filter to remove noisy points and Honolulu airport (outliers)

```{r}
#| eval: TRUE
#| echo: TRUE

#using |>
delays <- flights |>
  group_by(dest) |> 
  summarise(
    count = n(),
    dist = mean(distance, na.rm = TRUE),
    delay = mean(arr_delay, na.rm = TRUE)
  ) |>
  filter(count > 20, dest != "HNL")
```

## In-Class Activity

-   We are going to once again practice manipulating data in R

-   Get with a partner and work through [Week 4 \| Class Activity]{style="color:#000E2F"}

## For Next Class

-   Review what you learned today

-   Do new set of readings:

    -   We are going to examine causality
    -   [ QSS Chapter 2.1 - 2.4]{style="color:#000E2F"}
        -   Follow along the exercises with your own computer

-   Bring your laptops to class
