---
title: "Quiz 5"
author: "Your Name"
format: 
  html:
    embed-resources: true
    theme: cosomo
editor: visual
---

### Instructions

#### How to **complete** this assignment.

-   Attempt each exercise in order.

-   In each code chunk, if you see "<font color="789922">\# INSERT CODE HERE"</font>, then you are expected to add some code to create the intended output (Make sure to erase "\# INSERT CODE HERE" and place your code in its place).

-   If my instructions say to "Run the code below..." then you do not need to add any code to the chunk.

-   Many exercises may require you to type some text below the code chunk, interpreting the output and answering the questions.

-   Please follow the rules from the course syllabus regarding seeking help with this assignment.

#### How to **submit** this assignment.

-   When you are finished, click the "Render" button at the top of this panel. If there are no errors, an HTML file should pop up after a few seconds.

-   Take a look at the resulting HTML file that pops up. Make sure everything looks correct, your name is listed at the top, and that there is no 'junk' code or output.

-   Save the HTML file (to your local computer, and/or to a cloud location) as: **Quiz 5 "Insert Your Name"**.

-   Before you upload the file to HuskyCT, you will need to zip/compress the file.

-   Upload the problem set zipped/compressed HTML to HuskyCT. **Do not** upload the original .qmd version.

-   This assignment is <font color="darkred">**due Monday, April 28, 2025, no later than 4:00 pm Eastern**</font>. Points will be deducted for late submissions.

-   <font color="darkblue">TIP:</font> Start early so that you can troubleshoot any issues with knitting to HTML.

#### Grading Rubric

There are 8 possible points on this assignment.

**Baseline (C level work)**

-   Your .qmd file knits to HTML without errors.
-   You answer questions correctly but do not use complete sentences.
-   There are typos and 'junk code' throughout the document.
-   You do not put much thought or effort into the Reflection answers.

**Average (B level work)**

-   You use complete sentences to answer questions.
-   You attempt every exercise/question.

**Advanced (A level work)**

-   Your code is simple and concise.
-   Unnecessary messages from R are hidden from being displayed in the HTML.
-   Your document is typo-free.
-   At the discretion of the instructor, you give exceptionally thoughtful or insightful responses.

------------------------------------------------------------------------

##### **Exercise 1. (A) Download the appropriate libraries and packages (Hint: take special notice of the file extension for the `NC Counties` data (Hint: use `load()`)). (B) Using the `Covid April 14` and the `NC Counties` data, make a map of the cumulative number of confirmed cases, per capita, in each county in North Carolina. (3 points)**

```{r}
#insert code here
```

##### **Exercise 2. (A) Download the `Covid Data 1`. (B) Combine the `Covid Data 1` to the data frame that you created in Exercise 1. Next, create a variable that represents the number of confirmed cases that occurred from April 13 and April 14. Lastly, compute the confirmed per capita cases and visualize it in a map. (3 points)**

```{r}
#insert code here
```

<font color="darkred">**ANSWER:**</font>

##### **Exercise 3. (A) Utilize the `Covid April 14` and `US` data sets, make a map of the cumulative number of confirmed cases per capita, in each state in the United States. (2 points)**

```{r}
#insert code here
```

<font color="darkred">**ANSWER:**</font>
