---
title: "DSDA 1995"
subtitle: "Data Science and Society using R"
author: "Lecture 3 | R Packages and Getting/Loading Data"
format: 
  revealjs:
    slide-number: c/t
    scrollable: true
    theme: [simple, custom.scss]
    embed-resources: true
editor: visual
---

## Plan for Today

-   Examine R Packages

-   Importing data from different sources

## R Packages

-   R is an open-source programming language, meaning that users can contribute packages that make our lives easier, and we can use them for free. For today, and the future, we will use many R packages including:

-   The suite of **tidyverse** packages: for data wrangling and data visualization

-   **gapminder**: for easy access to an excerpt of the Gapminder data on life expectancy, GDP per capita, and population by country

## Install R Packages

-   If these packages are not already available in your R environment, install them by typing the following three lines of code into the console of your RStudio session. Note that you can check to see which packages (and which versions) are installed by inspecting the Packages tab in the lower right panel of RStudio.

<br>

```{r}
#| message: FALSE
#| eval: FALSE
#| echo: TRUE

install.packages("tidyverse")
install.packages("gapminder")
```

## Load R Library

-   You may need to select a server from which to download; any of them will work. Next, you need to load these packages in your working environment. We do this with the <code>library</code> function. Run the following lines in your console.

<br>

```{r}
#| message: FALSE
#| echo: TRUE

library(tidyverse)
library(gapminder)
```

------------------------------------------------------------------------

-   You only need to **install** packages once, but you need to **load** them each time you relaunch RStudio.

-   The Tidyverse packages share common philosophies and are designed to work together. You can find more about the packages in the tidyverse at <https://www.tidyverse.org>.

## Exploring the Gapminder Data

To get started, run the following command to load the data and save it to a local version called<code> gm </code>.

<br>

```{r}
#| echo: TRUE

gm <- gapminder::gapminder
```

<br>

-   This command instructs R to load some data: the gapminder data set. You should see that the environment area in the upper right hand corner of the RStudio window now lists a data set called gm that has 1704 observations of 6 variables. As you interact with R, you will create a series of objects. Sometimes you load them as we have done here, and sometimes you create them yourself as the byproduct of a computation or some analysis you have performed.

------------------------------------------------------------------------

-   We can view part of the data set by typing its name into the console.

<br>

```{r}
#| echo: TRUE

gm
```

------------------------------------------------------------------------

-   However, printing the whole dataset in the console is not that useful.

-   One advantage of RStudio is that it comes with a built-in data viewer. Click on the name gm in the Environment pane (upper right window) that lists the objects in your environment. This will bring up an alternative display of the data set in the Data Viewer (upper left window). You can close the data viewer by clicking on the x in the upper left hand corner.

-   What you should see are seven columns of numbers, each row representing a different combination of country and year: the first entry in each row is simply the row number (an index we can use to access the data from individual years if we want), the second is the country, followed by the continent in which the country is located, year, and the last three columns represent the life expectancy, population, and gross domestic product (GDP) per capita for that country in the given year, respectively. Use the scroll bar on the right side of the console window to examine the complete data set.

------------------------------------------------------------------------

-   Note that the row numbers in the first column are not part of the data. R adds them as part of its printout to help you make visual comparisons. You can think of them as the index that you see on the left side of a spreadsheet. In fact, the comparison to a spreadsheet will generally be helpful. R has stored this data set in a kind of spreadsheet or table called a data frame.

-   You can see the dimensions of this data frame as well as the names of the variables and the first few observations by typing:

<br>

```{r}
#| echo: TRUE

glimpse(gm)
```

<br>

or

<br>

```{r}
#| echo: TRUE

head(gm)
```

------------------------------------------------------------------------

-   It is better practice to type this command into your console, since it is not necessary code to include in your solution file.

-   We can see that there are 1704 observations and 6 variables in this dataset. The variable names are country, continent, year, lifeExp, pop, and gdpPercap.

## Further Exploration

-   Let’s start to examine the data a little more closely. We can access the data in a single column of a data frame separately using a command like

<br>

```{r}
#| eval: FALSE
#| echo: TRUE

gm$continent
```

<br>

-   This command will only show the names of continents given in each row of the data set. The dollar sign basically says “go to the data frame that comes before me, and find the variable that comes after me”.

1). What command would you use to extract just the country names?

-   Notice that the way R has printed these data is different. When we looked at the complete data frame, we saw 1704 rows, one on each line of the display. These data are no longer structured in a table with other variables, so they are displayed one right after another. Objects that print out in this way are called vectors; they represent a set of numbers. R has added numbers in \[brackets\] along the left side of the printout to indicate locations within the vector. For example, Asia follows \[1\], indicating that Asia is the first entry in the vector. And if \[43\] starts a line, then that would mean the first item on that line would represent the 43rd entry in the vector.

## Brief Introduction to Data Visualization

-   R has some powerful functions for making graphics. We can create a simple plot of the life expectancy of each country per year with the command

<br>

```{r}
#| echo: TRUE
#| out-width: 100%
#| out-height: 80%

ggplot(data = gm, aes(x = lifeExp, y = gdpPercap)) + 
  geom_point()
```

## ggplot()

-   We use the <code>ggplot()</code> function to build plots. If you run the plotting code in your console, you should see the plot appear under the Plots tab of the lower right panel of RStudio. Notice that the command above again looks like a function, this time with arguments separated by commas.

-   with <code>ggplot()</code>

    -   The first argument is always the dataset
    -   Next, you provide the variables from the dataset to be assigned to <code>aes</code>thetic elements of the plot, e.g. the x and the y axes.
    -   Finally, you use another layer, separated by a <code>+</code> to specify the geometric object for the plot. Since we want to scatterplot, we use <code>geom_point()</code>

------------------------------------------------------------------------

-   For instance, if you wanted to visualize the above plot using a line graph, you would replace <code>geom_point()</code> with <code>geom_line()</code>.

<br>

```{r}
#| echo: TRUE
#| out-width: 100%
#| out-height: 80%

gm1 <- gm |>
  group_by(year) |>
  summarize(avglifeExp = mean(lifeExp))

ggplot(data = gm1, aes(x = year, y = avglifeExp)) + 
  geom_line()
```

------------------------------------------------------------------------

-   You might wonder how you are supposed to know the syntax for the <code>ggplot</code> function. Thankfully, R documents all of its functions extensively. To learn what a function does and its arguments that are available to you, just type in a question mark followed by the name of the function that you’re interested in. Try the following in your console:

<br>

```{r}
#| tidy: FALSE
#| echo: TRUE

?ggplot
```

<br>

-   Notice that the help file replaces the plot in the lower right panel. You can toggle between plots and help files using the tabs at the top of that panel.

## Importing Data

-   Getting and loading data can be a difficult step
-   Location, Location, Location!
-   Where do your data live?
    -   In R
    -   On your computer (offline)
    -   Exists on the web
        -   Can be loaded view an R package
        -   Can be accessed via URL
        -   Can be downloaded to your computer and loaded using the file path

## From R Packages

```{r}
#| eval: TRUE
#| echo: TRUE

# From the "datasets" package, load in the AirPassengers data
datasets::AirPassengers
airpass <- datasets::AirPassengers
```

------------------------------------------------------------------------

```{r}
#| eval: TRUE
#| echo: TRUE

# From the "gapminder" package, load in the gapminder data
gapminder::gapminder
gm <- gapminder::gapminder
```

## From your Computer

-   Where is R currently looking for files on your machine?
    -   get working directory

<br>

```{r}
#| eval: FALSE
#| echo: TRUE

getwd()
```

<br>

-   Where do you want R to look for files?
    -   set working directory

<br>

```{r}
#| eval: FALSE
#| echo: TRUE

setwd("~/Desktop")
```

<br>

-   As we did above, you can set the working directory by simply typing your location into the console. Additionally, you can "point and click" within RStudio to set the working directory.

-   Session -\> Set Working Directory -\> "Choose your option"

-   Files -\> "Choose your file" -\> More -\> Set as Working Directory (In the lower right side window)

## Getting Data into R

-   How are the data stored?
-   What is the file extension (e.g., .csv, .xlsx)
-   Did the data import correctly?
-   Do we need to make any changes?

## How Are The Data Stored

-   Structured
    -   Quanitative
        -   Spreadsheets
-   Unstructured
    -   Qualitative
        -   Social Media Post, Emails, Business Reports
-   Temporal
    -   Stock Ticks
-   Geolocation (shapefiles)
    -   Georeferenced points or areas

## File Extensions

-   .csv (Comma Seperated Values)
-   .xlsx (Microsoft Excel Spreadsheet)\
-   .txt (Text File (What is the delimiter?))

## We Will Use...

-   Structured data from R packages or clean(ish) .csv files
-   Plain text rectangular files (or flat files)
-   Using functions from `tidyverse` to load the files
-   Spatial data files

## Helpful Tips

-   .csv is preferred
    -   If you have an Excel spreadsheet, save it as .csv
-   Look at the original data file
    -   Are there blank rows or columns?
    -   Is there metadata contained in the spreadsheet?
    -   Is there inconsistent/messy data?
    -   Do dates look like dates?
    -   How much data cleaning or wrangling is required?

## Break
