---
title: "DSDA 1995"
subtitle: "Data Science and Society using R"
author: "Lecture 10 | Exploring Data II"
format: 
  revealjs:
    slide-number: c/t
    scrollable: true
    theme: [simple, custom.scss]
    embed-resources: true
editor: visual
---

```{r}
#| include: FALSE
library(tidyverse)
knitr::opts_chunk$set(eval = TRUE, echo = TRUE)
```

## Plan for Today

-   Continue to explore data in R


## Covariation

::: extrapad
- Describes the behavior between variables
:::

::: extrapad
- Is the tendency for the values of two or more variables to vary together in a related way
:::


## Categorical and Continuous Variable

::: extrapad
- Let's say we want to explore the distribution of a continuous variable broken down by a categorical variable
:::

::: extrapad
- We could use `geom_freqpoly()`
:::

```{r}
#| eval: TRUE
#| echo: TRUE
#| out-width: 100%
#| out-height: 100%

ggplot(data = diamonds, mapping = aes(x = price)) + 
  geom_freqpoly(mapping = aes(colour = cut), binwidth = 500)
```


## Categorical and Continuous Variable

::: extrapad
- Issues
  + It is based on the raw counts, which means we cannot see variation for categories that have smaller numbers
:::  

::: extrapad  
- Let's examine the density instead
  + Which is the count standardized so that the area under each frequency polygon is one.
:::

```{r}
#| eval: TRUE
#| echo: TRUE
#| out-width: 100%
#| out-height: 100%

ggplot(data = diamonds, mapping = aes(x = price, y = ..density..)) + 
  geom_freqpoly(mapping = aes(colour = cut), binwidth = 500)
```

::: extrapad
- What do we notice? 
:::

::: extrapad
- Do we like this plot? Is it useful? 
:::

::: extrapad
- Any issues? 
:::


## Boxplots

::: extrapad
- Lets try an alternative, the `geom_boxplot()`
:::

```{r}
#| eval: TRUE
#| echo: TRUE
#| out-width: 100%
#| out-height: 100%

ggplot(data = diamonds, mapping = aes(x = cut, y = price)) +
  geom_boxplot()
```


## Boxplots

::: extrapad
- What are box plots? 
  + The box provides the IQR (25th to the 75th percentiles)
  + The bar, inside the box, provides us the median (50th percentile)
  + The whiskers extend from each end of the box and go to the farthest non-outlier point in the distribution
  + The points display observations that fall more than 1.5 X IQR from either side of the box
:::

::: extrapad
- What do we think about the boxplot? 
:::


## Categorical Variable Ordering

::: extrapad
- What if our categorical variable does not have an intrinsic order? 
:::

::: extrapad
- We can utilize the `reorder()` function
:::

```{r}
#| eval: TRUE
#| echo: TRUE
#| out-width: 100%
#| out-height: 100%

ggplot(data = mpg, mapping = aes(x = class, y = hwy)) +
  geom_boxplot()
```

<br>

```{r}
#| eval: TRUE
#| echo: TRUE
#| out-width: 100%
#| out-height: 100%

ggplot(data = mpg) +
  geom_boxplot(mapping = aes(x = reorder(class, hwy, FUN = median), y = hwy))
```

<br>

```{r}
#| eval: TRUE
#| echo: TRUE
#| out-width: 100%
#| out-height: 100%

ggplot(data = mpg) +
  geom_boxplot(mapping = aes(x = reorder(class, hwy, FUN = median), y = hwy)) +
  coord_flip()
```


## Two Categorical Variables

::: extrapad
- To visualize the covariation between categorical variables, you’ll need to count the number of observations for each combination
:::

```{r}
#| eval: TRUE
#| echo: TRUE
#| out-width: 100%
#| out-height: 100%

ggplot(data = diamonds) +
  geom_count(mapping = aes(x = cut, y = color))
```

<br>

- The size of each circle in the plot displays how many observations occurred at each combination of values

## Two Categorical Variables

::: extrapad
- We can also utilize `geom_tile()`
:::

```{r}
#| eval: TRUE
#| echo: TRUE
#| out-width: 100%
#| out-height: 100%

diamonds %>% 
  count(color, cut)

diamonds %>% 
  count(color, cut) %>%  
  ggplot(mapping = aes(x = color, y = cut)) +
    geom_tile(mapping = aes(fill = n))
```


## Two Continuous Variables

::: extrapad
- The scatterplot
:::

```{r}
#| eval: TRUE
#| echo: TRUE
#| out-width: 100%
#| out-height: 100%

ggplot(data = diamonds) +
  geom_point(mapping = aes(x = carat, y = price))
```


## Two Continuous Variables

::: extrapad
- What happens as the number of observations grow in a scatterplot? 
:::

::: extrapad
- Using the alpha aesthetic to add transparency
:::

::: extrapad
- Another option is to use `geom_bin2d()` and `geom_hex()`, which divide the coordinate plane into 2d bins and then use a fill color to display how many points fall into each bin
:::

```{r}
#| eval: TRUE
#| echo: TRUE
#| out-width: 100%
#| out-height: 100%

ggplot(data = diamonds) + 
  geom_point(mapping = aes(x = carat, y = price), alpha = 1 / 100)

ggplot(data = diamonds) +
  geom_bin2d(mapping = aes(x = carat, y = price))

# install.packages("hexbin")
ggplot(data = diamonds) +
  geom_hex(mapping = aes(x = carat, y = price))
```

## Two Continuous Variables

::: extrapad
- Additionally, we can bin one of the continuous variables so that it acts like a categorical variable
:::

::: extrapad
- `cut_width(x, width)`, as used below, divides x into bins of width width
:::

```{r}
#| eval: TRUE
#| echo: TRUE
#| out-width: 100%
#| out-height: 100%

ggplot(data = diamonds, mapping = aes(x = carat, y = price)) + 
  geom_boxplot(mapping = aes(group = cut_width(carat, 0.1)))
```

::: extrapad
- Another approach is to display approximately the same number of points in each bin using `cut_number()`
:::

```{r}
#| eval: TRUE
#| echo: TRUE
#| out-width: 100%
#| out-height: 100%

ggplot(data = diamonds, mapping = aes(x = carat, y = price)) + 
  geom_boxplot(mapping = aes(group = cut_number(carat, 20)))
```



## Patterns and Models

::: extrapad
- Patterns
  - Patterns in your data provide clues about relationships
:::

::: extrapad
- If a systematic relationship exists between two variables it will appear as a pattern in the data
  + Could this pattern be due to coincidence (i.e. random chance)?
  + How can you describe the relationship implied by the pattern?
  + How strong is the relationship implied by the pattern?
  + What other variables might affect the relationship?
  + Does the relationship change if you look at individual subgroups of the data?
:::  

## Patterns and Models

::: extrapad
- A scatterplot of Old Faithful eruption lengths versus the wait time between eruptions shows a pattern
:::
```{r}
#| eval: TRUE
#| echo: TRUE
#| out-width: 100%
#| out-height: 100%

ggplot(data = faithful) + 
  geom_point(mapping = aes(x = eruptions, y = waiting))
```

<br>

- What do we notice? 

- We will discuss models later on in the semester. 


## For Next Class

-   Review what you learned today

-   Do new set of readings:

    -   We are going to examine causality
    -   [R4DS Chapter 5]{style="color:#000E2F"}
    -   [R4DS Chapter 19.1 - 19.2]{style="color:#000E2F"}
        -   Follow along the exercises with your own computer

-   Bring your laptops to class
