---
title: "Problem Set 5"
author: "Your Name"
format: 
  html:
    embed-resources: true
    theme: cosomo
editor: visual
---

### Instructions

#### How to **complete** this assignment.

-   Attempt each exercise in order.

-   In each code chunk, if you see "<font color="789922">\# INSERT CODE HERE"</font>, then you are expected to add some code to create the intended output (Make sure to erase "\# INSERT CODE HERE" and place your code in its place).

-   If my instructions say to "Run the code below..." then you do not need to add any code to the chunk.

-   Many exercises may require you to type some text below the code chunk, interpreting the output and answering the questions.

-   Please follow the rules from the course syllabus regarding seeking help with this assignment.

#### How to **submit** this assignment.

-   When you are finished, click the "Render" button at the top of this panel. If there are no errors, an HTML file should pop up after a few seconds.

-   Take a look at the resulting HTML file that pops up. Make sure everything looks correct, your name is listed at the top, and that there is no 'junk' code or output.

-   Save the HTML file (to your local computer, and/or to a cloud location) as: **Problem Set 5 "Insert Your Name"**.

-   Before you upload the file to HuskyCT, you will need to zip/compress the file.

-   Upload the problem set zipped/compressed HTML to HuskyCT. **Do not** upload the original .qmd version.

-   This assignment is <font color="darkred">**due Monday, April 21, 2025, no later than 4:00 pm Eastern**</font>. Points will be deducted for late submissions.

-   <font color="darkblue">TIP:</font> Start early so that you can troubleshoot any issues with knitting to HTML.

#### Grading Rubric

There are 8 possible points on this assignment.

**Baseline (C level work)**

-   Your .qmd file knits to HTML without errors.
-   You answer questions correctly but do not use complete sentences.
-   There are typos and 'junk code' throughout the document.
-   You do not put much thought or effort into the Reflection answers.

**Average (B level work)**

-   You use complete sentences to answer questions.
-   You attempt every exercise/question.

**Advanced (A level work)**

-   Your code is simple and concise.
-   Unnecessary messages from R are hidden from being displayed in the HTML.
-   Your document is typo-free.
-   At the discretion of the instructor, you give exceptionally thoughtful or insightful responses.

------------------------------------------------------------------------

##### **Exercise 1. (A) Read in the `chetty` data and load the appropriate packages. (B) Explain what the following code does. You may look up any new functions in R Studio or on the web to understand them. (2 points)**

```{r}
#A
library(tidyverse)

chetty <- read_csv("Chetty_2014.csv")

#B
chetty_state_level <- chetty %>%
  group_by(state) %>%
  summarize(totpop = sum(pop_2000),
            region = first(region),
            division = first(division),
            inctax_progressive = first(inctax_progressive),
            avg_mob = weighted.mean(abs_mobility, pop_2000, na.rm=TRUE))
```

<font color="darkred">**ANSWER:**</font>

The above code collapses the original data frame into one observation per state. Region, division and inctax_progressive are all the same for every CZ in the state, so this code takes the corresponding value of the first CZ. Population for the state is obtained by summing over all the CZs in the state, and average mmobility in the state is obtained by averaging over all CZs, weighted by the population in each one to get a better representation of the average across individuals in the state.

##### **Exercise 2. (A) Use the `crosswalk` data to add variables from `chetty_state_level` to the data in the `segregation` data. Select the appropriate variables so the end result is the data frame in `utip`, minus the variable inctax. (B) Create and add the variable inctax to the data frame to get exactly what we had in `utip`. (2 points)**

```{r}
#INSERT CODE HERE
```

<font color="darkred">**ANSWER:**</font>

##### **Exercise 3. Make a tidy version of `tuition` data. (2 points)**

```{r}
#INSERT CODE HERE
```

<font color="darkred">**ANSWER:**</font>

##### **Exercise 4. (A) What goes wrong when you try to join the data frame in `ccd` with the data frame in `school` by school name? Print out the rows that are matched and the rows that are not matched. (B) Make one simple modification to one of these data frames that dramatically increases the number of rows that are matched, and print out the rows that match and the rows that do not match after this modification. (2 points)**

```{r}
#INSERT CODE HERE
```

<font color="darkred">**ANSWER:**</font>
