---
title: "Week 2 | In-Class Activity"
format: 
  html:
    embed-resources: true
editor: visual
css: lab.css
---

## Class Activity

We are going to practice finding, loading, and dealing with messy data.

-   About the data
    -   Water consumption in NYC
    -   [NYC Open Data](https://opendata.cityofnewyork.us)

------------------------------------------------------------------------

### Getting the Data

-   Go to: [the course website](https://github.com/byersjs/POLS-5605)
    -   Download the data
    -   Load the `tidyverse` package

------------------------------------------------------------------------

### Utilize This Code with a Partner

```{r eval = FALSE}
#Figure out what working directory R is currently utilizing
getwd()
```

```{r eval = FALSE}
#Where did you download the data (i.e., desktop, download folder, etc.)
#Do you need to change your working directory to ensure that you can load the data?
setwd("insert your working directory here")
```

```{r eval = FALSE}
#Once you have set your working directory, make sure to read the data into R

nyc <- read.csv("NYCWater.csv")
```

```{r eval = FALSE}
glimpse(carib)
```

1.  Remove the first row from the data set.

```{r eval = FALSE}
#We are going to delete the first row of the data
#nyc[rownumber, columnnumber]

nyc <- nyc[-1, ]

#

#Check to see if it worked
View(nyc)
```

Do we notice any issues with the data set?

```{r eval = FALSE}
View(nyc)
```

The first thing you should notice is that all of the variable types are numeric. Additionally, all of the variable names are too long and complicated.

```{r eval = FALSE}
class(nyc$New.York.City.Population)
```

1.  What does R mean when it says a variable (in this case the column for "Island Names") is a numeric class?

2.  What is the code below showing you?

```{r eval = FALSE}
nyc$Year[1:5]
```

1.  Which years are represented in this dataset?

```{r eval = FALSE}
unique(nyc$Year)
```

Let's create a plot

```{r eval = FALSE}
ggplot(data = nyc, mapping = aes(x = year, y = Per.Capita.Gallons.per.person.per.day.)) + 
  geom_point()
```

1.  Rename the "New.York.City.Population" and "NYC.Consumption.Million.gallons.per.day." variables

```{r eval = FALSE}
colnames(dataFrame)[colnames(dataFrame) == "oldName"] <- "newName"

# OR

dataframe <- dataframe |>
  rename(NewName = Oldname)
```
