---
title: " | In-Class Activity"
format: 
  html:
    embed-resources: true
editor: visual
css: custom.css
runtime: shiny
---

```{r}
#| include: FALSE

library(tidyverse)

```

## Plan for Today

- Explore Shiny apps in R
- Explore basic UI
- Explore basic reactivity


---

## Shiny App Introduction

- What is Shiny?

  + Shiny is a framework for creating web applications using R code
  
  + It is designed primarily with data scientists in mind, and to that end, you can create pretty complicated Shiny apps with no knowledge of HTML, CSS, or JavaScript
  
  + On the other hand, Shiny doesn’t limit you to creating trivial or prefabricated apps: its user interface components can be easily customized or extended, and its server uses reactive programming to let you create any type of back end logic you want
  
- What do people use Shiny for?

  + Create dashboards that track important high-level performance indicators, while facilitating drill down into metrics that need more investigation
  
  + Replace hundreds of pages of PDFs with interactive apps that allow the user to jump to the exact slice of the results that they care about
  
  + Communicate complex models to a non-technical audience with informative visualizations and interactive sensitivity analysis
  

## Shiny App

- Let's install `shiny` and start putting together an app

```{r}
#install.packages("shiny")

library(shiny)
```


## Shiny App

- There are several ways to create a Shiny app. The simplest is to create a new directory for your app, and put a single file called `app.R` in it

  + This `app.R` file will be used to tell Shiny both how your app should look, and how it should behave
  
```{r}

library(shiny)

selectInput("dataset", label = "Dataset", choices = ls("package:datasets"))

```

```{r}
renderPrint({
  req(input$dataset)
  data <- get(input$dataset, "package:datasets")
  summary(data)
})
```






