---
title: " | In-Class Activity"
format: 
  html:
    embed-resources: true
editor: visual
css: custom.css
runtime: shiny
---


---

## title: "Intro to Shiny with Quarto" format: revealjs editor: visual execute: echo: true warning: false message: false runtime: shiny

## Setup

We'll use the `shiny` package to build interactive apps.

```{r}
library(shiny)
```

---

## A Minimal App

```{r}
"Hello, world!"
```

Quarto doesn't require `fluidPage()` or `shinyApp()` — your document is the UI. Code chunks containing `render*()` functions are reactive and act like Shiny server code.

---

## Dataset Viewer

### Choose a Dataset

```{r}
selectInput("dataset", "Dataset", choices = ls("package:datasets"))
```

---

### Summary of Selected Dataset

```{r}
renderPrint({
  dataset <- get(input$dataset, "package:datasets")
  summary(dataset)
})
```

---

### First Six Rows

```{r}
renderTable({
  dataset <- get(input$dataset, "package:datasets")
  head(dataset)
})
```

---

## Inputs: Text and Numbers

```{r}
textInput("name", "What's your name?")
numericInput("age", "Your Age", value = 30)
```

---

## Output Greeting

```{r}
renderText({
  paste("Hello,", input$name, "— Age:", input$age)
})
```

---

## Basic Plotting

### Simple Base R Plot

```{r}
renderPlot({
  plot(1:10)
})
```

---

### ggplot2 Plot

```{r}
library(ggplot2)

renderPlot({
  ggplot(mpg) +
    geom_point(aes(displ, hwy))
})
```

---

## Basic Reactivity

```{r}
textInput("user", "Enter your name:")

renderText({
  paste("Hello", input$user)
})
```

Reactive expressions update automatically when inputs change.

---

## Layout with Columns

\::: columns ::: column

```{r}
sliderInput("n", "Number of obs", min = 10, max = 1000, value = 100)
```

\:::

\::: column

```{r}
renderPlot({
  hist(rnorm(input$n))
})
```

\::: :::

---

## Simulate and Analyze

```{r}
numericInput("n1", "Sample Size 1", value = 100)
numericInput("n2", "Sample Size 2", value = 100)

renderText({
  x1 <- rnorm(input$n1)
  x2 <- rnorm(input$n2)
  t.test(x1, x2)$p.value
})
```

---

## Tidy Evaluation in Shiny + Quarto

```{r}
library(tidyverse)

selectInput("var", "Variable", choices = c("carat", "price"))
numericInput("min", "Minimum", value = 1)

renderTable({
  diamonds %>%
    filter(.data[[input$var]] > .env$input$min) %>%
    head()
})
```

---

## Summary: Quarto vs Classic Shiny

| Feature                | Rmd + shinyApp()            | Quarto + `runtime: shiny`         |
| ---------------------- | --------------------------- | --------------------------------- |
| UI Layout              | `fluidPage()`               | Markdown sections + layout chunks |
| Inputs/Outputs         | `selectInput`, `plotOutput` | Used directly in chunks           |
| `shinyApp(ui, server)` | Required                    | ❌ Not needed                      |
| Render Functions       | In `server` block           | In R code chunks with `render*()` |
| Layout                 | `fluidRow()`, `column()`    | `columns`, `panel-tabset`, etc.   |

---

## Next Steps

- Add sidebar panels with `#| panel: sidebar`
- Use `reactive()` expressions to optimize reactivity
- Add interactivity with `actionButton()` and `observeEvent()`

Happy Shiny building with Quarto! 🚀

